---
title: "SaveSpotHomework2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2. Building functions

Let's first create our data set and instantiate the proper variables. We use a modified version of the function from question 1 to build our dummy data and solve for the standard errors various ways.

```{r, echo=F}
set.seed(42)

# First we build our regression function

reg_test = function(rep, omega, index){
  # Instantiate vector of zeros 
  n = rep(0,100)
  n = replace_func(index, n) # Use the replace function to build the dummy vector
  p = length(which(n==1))
  # Build error vector based on inputs omega and index
  errors = c(rnorm(p, 0, omega^2), rnorm(length(n)-p, 0, 1))
  
  df = tibble(xx = n, # Dummy variables for data generation are repressed by beta = 0
              err = errors,
              y = err)
  
  reg = lm(y~xx, data=df)
  sigmas <- c(rep(omega, p), rep(1, length(n)-p))
  results <-
    list(
      beta = coefficients(reg)[2],
      sigB_0 = sqrt(vcovHC(reg, type = 'const')[2,2]),
      sigB_ideal = sqrt(vcovHC(reg, omega = sigmas ^ 2)[2,2]),
      sigB_HC0 = sqrt(vcovHC(reg, type ='HC0')[2,2]),
      sigB_HC1 = sqrt(vcovHC(reg, type ='HC1')[2,2]),
      sigB_HC2 = sqrt(vcovHC(reg, type ='HC2')[2,2]),
      tstat = coef(summary(reg))["xx", "t value"]
    )  # extract the slope coefficient from each simulation
  return(results)
}

## Parameter Settings
param_list.2 <- expand.grid(rep=1:100, omega = c(1,2),
                            index=c(3,10,25))

## Run simulation
sim_out.2 <- param_list.2 %>%
  mutate(results=pmap(param_list.2,reg_test)) %>%
  unnest_wider(results)

```

## 2. Results

Below is the simulation output for our standard errors given each HCx type and each combination of $\omega$ and $n_1$. We notice that our HCx groups give much higher standard error values for $\omega = 2$ for low values of $n_1$ and slowly shrink as $n_1$ grows. This is to be expected as we also note the T-stat showing relatively high probability of rejection for $\omega=2$ which shrinks as $n_1$ grows. This shows that as we approach a balanced dummy variable vector that the need for corrections from HC goes away.

When $n_1 = 25$ we find that nearly all of our $\omega=1$ estimates agree, and that the $\omega=2$ estimates show our ideal and 0 correction terms closely related where the HC group is much higher, again confirming our findings.

```{r}
## Print result
sim_print <- group_by(sim_out.2,index, omega) %>%
  mutate(sd_beta2=sd(beta), mean_0=mean(sigB_0), mean_idl=mean(sigB_ideal), mean_HC0=mean(sigB_HC0), mean_HC1=mean(sigB_HC1), mean_HC2 = mean(sigB_HC2), p_rej= length(tstat[tstat>1.96])/length(tstat)) %>%
  ungroup() %>%
  select(index,omega,sd_beta2, mean_0, mean_idl, mean_HC0, mean_HC1, mean_HC2, p_rej) %>%
  distinct() 

sim_print
```



#############################################

Save spot
